\section{Introduction}\label{main_Introduction}
The capacitated vehicle routing problem with time windows or CVRP-TW is a combinatorial optimization problem seeking to service a number of customers with a fleet of vehicles where the vehicles have a mimited capacity and the delivery locations have time windows within which the deliveries (or visits) must be made. Often the context is that of delivering goods located at a central depot to customers who have placed orders for such goods. Implicit is the goal of minimizing the cost of distributing the goods. Finding the global minimum for the cost function, except for the smallest instances, is computationally complex.\section{AUTHORS}\label{main_authors}
\begin{TabularC}{3}
\hline
Dolphin project-team INRIA Futurs, 2007. &Antonio La\-Torre atorre[at]fi.upm.es &Thomas Legrand paradiseo-help[at]lists.gforge.inria.fr  \\\hline
\end{TabularC}
\section{LICENSE}\label{main_LICENSE}
This software is governed by the Ce\-CILL license under French law and abiding by the rules of distribution of free software. You can use, modify and/ or redistribute the software under the terms of the Ce\-CILL license as circulated by CEA, CNRS and INRIA at the following URL \char`\"{}http://www.cecill.info\char`\"{}.

As a counterpart to the access to the source code and rights to copy, modify and redistribute granted by the license, users are provided only with a limited warranty and the software's author, the holder of the economic rights, and the successive licensors have only limited liability.

In this respect, the user's attention is drawn to the risks associated with loading, using, modifying and/or developing or reproducing the software by the user in light of its specific status of free software, that may mean that it is complicated to manipulate, and that also therefore means that it is reserved for developers and experienced professionals having in-depth computer knowledge. Users are therefore encouraged to load and test the software's suitability as regards their requirements in conditions enabling the security of their systems and/or data to be ensured and, more generally, to use and operate it in the same conditions as regards security. The fact that you are presently reading this means that you have had knowledge of the Ce\-CILL license and that you accept its terms.

Paradis\-EO Web\-Site : \tt{http://paradiseo.gforge.inria.fr} Contact: \tt{paradiseo-help@lists.gforge.inria.fr}\section{Home Page}\label{main_Paradiseo}
{\tt http://paradiseo.gforge.inria.fr} 